### 2025-04-03 _archive/ellipsis-projection.qmd
Initial posterior work done, missing causal section

---
title: Ellipsis Projection
author: Daniel Posmik
date: today
format: pdf
toc: false
number_sections: true
---

```{r setup, include=FALSE}
# Set up knit environment
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(error = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```

## Simulation of Points on an Elliptic Curve

First, we set up our parameters and generate points on an elliptic curve.

```{r}
# Set random seed for reproducibility
set.seed(123)

# Parameters for the elliptic curve
a <- 3  # semi-major axis
b <- 1  # semi-minor axis

# Generate 5 points on an elliptic curve
n <- 20
theta <- seq(0, 2*pi, length.out = n+1)[1:n]  # equally spaced angles

# Generate points on the perfect ellipse
x_ellipse <- a * cos(theta)
y_ellipse <- b * sin(theta)

# Add Gaussian noise N(0,1) to create observed points
noise_sd <- 0.2
x_observed <- x_ellipse + rnorm(n, 0, noise_sd)
y_observed <- y_ellipse + rnorm(n, 0, noise_sd)

# Generate uniform(0,1) values for the z coordinate
z_values <- runif(n)

# Create the data frame
data <- data.frame(
  x = x_observed,
  y = y_observed,
  z = z_values
)
```

## Projection onto the Ellipse

Here we project the noisy points back onto the ellipse, considering all three dimensions.

```{r}
# Project a point onto the ellipse and return coordinates and angles
project_to_ellipse <- function(x, y, z, a, b) {
  # Starting guess for theta (x-y plane angle)
  theta_guess <- atan2(y/b, x/a)

  # Function to minimize: squared distance from (x,y,z) to a point on the ellipse
  distance_function <- function(theta) {
    ellipse_x <- a * cos(theta)
    ellipse_y <- b * sin(theta)
    ellipse_z <- 0 # Since our ellipse only lives in x,y space
    return(sqrt((x - ellipse_x)^2 + (y - ellipse_y)^2 + (z - ellipse_z)^2))
  }

  # Optimize to find the best theta
  result <- optimize(distance_function, c(0, 2*pi))
  theta_optimal <- result$minimum

  # Calculate projected coordinates
  x_proj <- a * cos(theta_optimal)
  y_proj <- b * sin(theta_optimal)
  z_proj <- 0

  # Calculate distance between original and projected point
  dist <- sqrt((x - x_proj)^2 + (y - y_proj)^2 + (z - z_proj)^2)

  # Calculate the second angle (phi) - elevation from x-y plane
  # This is the angle between the x-y plane and the line to the original point
  phi <- atan2(z, sqrt(x^2 + y^2))

  # Return the projected point coordinates, angles, and distance
  return(list(
    x_proj = x_proj,
    y_proj = y_proj,
    z_proj = z_proj,
    theta = theta_optimal,  # Azimuthal angle in x-y plane
    phi = phi,              # Elevation angle from x-y plane
    distance = dist
  ))
}

# Apply the projection to each point
projection_results <- lapply(1:n, function(i) {
  project_to_ellipse(data$x[i], data$y[i], data$z[i], a, b)
})

# Extract results into data frame
data$x_projected <- sapply(projection_results, function(res) res$x_proj)
data$y_projected <- sapply(projection_results, function(res) res$y_proj)
data$z_projected <- sapply(projection_results, function(res) res$z_proj)
data$projection_theta <- sapply(projection_results, function(res) res$theta)
data$projection_phi <- sapply(projection_results, function(res) res$phi)
data$projection_distance <- sapply(projection_results, function(res) res$distance)

# Print the data frame with projection information
head(data)
```

## 2D Visualization

```{r, fig.width=8, fig.height=8}
library(ggplot2)

# Create a dense set of points for drawing the perfect ellipse
theta_dense <- seq(0, 2*pi, length.out = 100)
ellipse_points <- data.frame(
  x = a * cos(theta_dense),
  y = b * sin(theta_dense)
)

# Create the plot
p <- ggplot() +
  # Draw the perfect ellipse
  geom_path(data = ellipse_points, aes(x = x, y = y), color = "gray50", linetype = "dashed") +
  # Draw the observed points with color based on z-value
  geom_point(data = data, aes(x = x, y = y, color = z), size = 3) +
  # Add a color scale for the z-values
  scale_color_viridis_c(name = "Z Value") +
  # Draw the projected points
  geom_point(data = data, aes(x = x_projected, y = y_projected), color = "purple", shape = 1, size = 3) +
  # Draw lines connecting observed points to their projections
  geom_segment(data = data, 
               aes(x = x, y = y, xend = x_projected, yend = y_projected),
               linetype = "dotted") +
  # Add labels and theme
  labs(title = "Elliptic Curve Simulation and Projection (X-Y Plane View)",
       subtitle = "Original points (colored by z-value) and their projections (purple circles)",
       x = "X", y = "Y") +
  theme_minimal() +
  coord_equal()  # Equal aspect ratio for proper ellipse visualization

# Display the plot
print(p)
```

## Summary of Projection Metrics

```{r}
# Create summary table
summary_table <- data.frame(
  Point = 1:n,
  Original_X = data$x,
  Original_Y = data$y,
  Original_Z = data$z,
  Projected_X = data$x_projected,
  Projected_Y = data$y_projected,
  Projected_Z = data$z_projected,
  Theta_Angle = data$projection_theta,  # Azimuthal angle in x-y plane
  Phi_Angle = data$projection_phi,      # Elevation angle from x-y plane
  Projection_Distance = data$projection_distance
)

# Print the summary table
head(summary_table)
```

# **Code Appendix**

```{r, ref.label = knitr::all_labels()}
#| echo: true
#| eval: false
```